### 2024-08-01 Paper/appendix.tex
Updates from Overleaf
Update tex

\subsection{Multiverse of the effect in different comparison groups and parameter choices, 1901} 
\FloatBarrier
\begin{table}[H]
\centering
\caption{Regression results for population size} \label{tab:pop1}
\footnotesize
\begin{tabular}{lcc}
   \tabularnewline \midrule \midrule
   Outcome: & \multicolumn{2}{c}{log(Population)}\\
                                                    & (1)             & (2)\\  
                                                    & Dummy approach  & Market access approach\\  
   \midrule
   Year 1787 $\times$ Affected                       & 0.0160          & 0.0379\\   
                                                     & (0.0106)        & (0.0626)\\   
   Year 1834 $\times$ Affected                       & 0.0177          & -0.0146\\   
                                                     & (0.0127)        & (0.0748)\\   
   Year 1840 $\times$ Affected                       & -0.0060         & -0.2651$^{***}$\\   
                                                     & (0.0132)        & (0.0759)\\   
   Year 1845 $\times$ Affected                       & -0.0029         & -0.2531$^{***}$\\   
                                                     & (0.0143)        & (0.0775)\\   
   Year 1850 $\times$ Affected                       & 0.0021          & -0.2466$^{***}$\\   
                                                     & (0.0152)        & (0.0900)\\   
   Year 1860 $\times$ Affected                       & 0.0319$^{*}$    & 0.0183\\   
                                                     & (0.0175)        & (0.0972)\\   
   Year 1880 $\times$ Affected                       & 0.1379$^{***}$  & 0.8349$^{***}$\\   
                                                     & (0.0203)        & (0.1155)\\   
   Year 1901 $\times$ Affected                       & 0.2364$^{***}$  & 1.591$^{***}$\\   
                                                     & (0.0264)        & (0.1607)\\  
   \midrule
   Observations                                     & 14,301          & 14,301\\
   Parishes                                         & 1,589           & 1,589\\
   \midrule 
   Parish FE                                        & Yes             & Yes\\  
   Year FE                                          & Yes             & Yes\\
   \midrule \midrule
\end{tabular}
\parbox{0.9\textwidth}{
\caption*{\textit{Notes:} Cluster-robust standard errors in the parenthesis. Clustered at the parish level. Affected is either a dummy for being in the West Limfjord or improvement in market access, which is indicated by the headers of the results. *** $p< 0.01$ ** $p< 0.05$ * $p< 0.10$. \\ \textit{Source: Danish census data}}
}
\end{table}

@@

    for b in 1 to B: # Loop of MC samples
    	# Generate samples from Y_max_c Y_max_c:
    	for c in 1 to C:
    		t_c = sample_uniform(1, Y_min_c, Y_max_c)
    		# Is t_c equal to t?
    		coins_t[c] = t_c == t

    	# Were there any coins associated with this time in this draw?
    	number_of_coins = sum(coins_t)
    	succces_t[b] = number_of_coins > 0

>>>>

\subsection{Full regression table} 
\FloatBarrier
\begin{table}[H]
\centering
\caption{Regression results for population size} \label{tab:pop1}
\footnotesize
\begin{tabular}{lcc}
   \tabularnewline \midrule \midrule
   Outcome: & \multicolumn{2}{c}{log(Population)}\\
                                                    & (1)             & (2)\\  
                                                    & Dummy approach  & Market access approach\\  
   \midrule
   Year 1787 $\times$ Affected                       & 0.0160          & 0.0379\\   
                                                     & (0.0106)        & (0.0626)\\   
   Year 1834 $\times$ Affected                       & 0.0177          & -0.0146\\   
                                                     & (0.0127)        & (0.0748)\\   
   Year 1840 $\times$ Affected                       & -0.0060         & -0.2651$^{***}$\\   
                                                     & (0.0132)        & (0.0759)\\   
   Year 1845 $\times$ Affected                       & -0.0029         & -0.2531$^{***}$\\   
                                                     & (0.0143)        & (0.0775)\\   
   Year 1850 $\times$ Affected                       & 0.0021          & -0.2466$^{***}$\\   
                                                     & (0.0152)        & (0.0900)\\   
   Year 1860 $\times$ Affected                       & 0.0319$^{*}$    & 0.0183\\   
                                                     & (0.0175)        & (0.0972)\\   
   Year 1880 $\times$ Affected                       & 0.1379$^{***}$  & 0.8349$^{***}$\\   
                                                     & (0.0203)        & (0.1155)\\   
   Year 1901 $\times$ Affected                       & 0.2364$^{***}$  & 1.591$^{***}$\\   
                                                     & (0.0264)        & (0.1607)\\  
   \midrule
   Observations                                     & 14,301          & 14,301\\
   Parishes                                         & 1,589           & 1,589\\
   \midrule 
   Parish FE                                        & Yes             & Yes\\  
   Year FE                                          & Yes             & Yes\\
   \midrule \midrule
\end{tabular}
\parbox{0.9\textwidth}{
\caption*{\textit{Notes:} Cluster-robust standard errors in the parenthesis. Clustered at the parish level. Affected is either a dummy for being in the West Limfjord or improvement in market access, which is indicated by the headers of the results. *** $p< 0.01$ ** $p< 0.05$ * $p< 0.10$. \\ \textit{Source: Danish census data}}
}
\end{table}

@@

    for b in 1 to B: # Loop of MC samples
    ...	# Generate samples from Y_max_c Y_max_c:
    ...	for c in 1 to C:
    ...	...	t_c = sample_uniform(1, Y_min_c, Y_max_c)
    ...	...	# Is t_c equal to t?
    ...	...	coins_t[c] = t_c == t

    ...	# Were there any coins associated with this time in this draw?
    ...	number_of_coins = sum(coins_t)
    ...	succces_t[b] = number_of_coins > 0

### 2024-08-01 Paper/main.tex
Update tex

    \normalsize
    \begin{changemargin}{1cm}{1cm}
    Is geography destiny? What is the role of first-nature geography in determining prosperity? This paper estimates the effect of randomly removing and introducing favorable first-nature geography to a specific region using a difference in difference design. In 1825 a storm created a new natural navigable waterway, bringing trade and prosperity to the otherwise relatively isolated northwestern Denmark. 700 years prior, the same event happened in reverse, when a previous channel closed up between 1086 and 1208. The elasticity of geography-induced market access is estimated to be 1.6, corresponding to 24 percent population growth within a generation of the event. Demonstrated mechanisms include trade, fertility, fishing, and the rise of manufacturing. The central finding is replicated in reverse in a register of dated archaeological sites. The 1086-1208 closing caused fewer buildings and sites containing coins. The general insight is the same: First-nature geography determines the levels and location of prosperity.

@@

This paper provides evidence of the effects of these shocks in a difference-in-difference framework. This is done using historical census data, a large database of Northern European historical trade (the Sound Toll Register), and the densities of archaeological findings. It is shown that the 1825 breach led to an increase in population of 24 percent in 1901. To document the mechanisms, I have collected new data on all Danish occupations (13.5 million observations) in the period 1787-1901. The Danish censuses contain a description of each person's occupation. A new machine learning procedure, \textit{OccCANINE}, is applied to these descriptions to convert them into standardized HISCO codes with high accuracy \citep{leeuwen2002hisco, dahl2024breaking}. The HISCO-coded census data is then used to estimate the effect on occupational structure. The key finding is that the channel caused both fishing and manufacturing to grow. An important question is whether the improved geography caused growth or simply reallocation of prosperity. By using age, gender, and birthplace information it is shown that the population growth came primarily from fertility - not migration. Denmark in the 19th century was in a post-Malthusian regime, where improved living standards cause fertility \citep{Jensen2022, Galor2011}. Taken together, the evidence of this paper suggests that intrinsic growth is the most likely explanation of the observed effect. That is, improved first-nature geography seemingly causes intrinsic growth - not just a reallocation of growth that would have occurred elsewhere.

@@

The improved first-nature geography also caused the population to increase. Figure \ref{fig:pop1} shows regression results from estimating equation \ref{eq:eq501}. Panel (a) shows the simplest estimate, where being affected by the Agger channel is defined as a dummy for being in the West Limfjord. This allows parameters to be interpreted as the approximate percentage increase in the affected region caused by the channel. By 1901 the parishes of the West Limfjord region had a 23.6 percent larger population than what would have been expected by parishes from elsewhere and the population in 1801. Panel (b) shows the results when 'affected' is defined as estimated improved market access from the channel. This shows that by 1901, affected areas had experienced a population growth equivalent to an elasticity of market access of 1.59.\footnote{This is similar to the magnitude found for the Panama Canal by \cite{rauch2022a}.}

@@

The empirical findings of this study reveal that the sudden emergence of the Agger channel in 1834 had a profound and lasting impact on the economic landscape of Denmark's West Limfjord region. Using a difference-in-differences design, it is shown that the channel's creation led to a 24 percent increase in the population of affected parishes by 1901 compared to unaffected areas. This population growth was driven primarily by enhanced market access, which facilitated a significant rise in trade, fishing and manufacturing activities. Moreover, the findings indicate that this growth was intrinsic, with higher fertility rates contributing to the increased population rather than migration. Additionally, a reverse natural experiment from the 12th century, when a similar channel closed, resulted in a substantial decline in economic activity, as evidenced by archaeological data. These results underscore the critical role of first-nature geography in driving economic prosperity.

>>>>

    \normalsize
    \begin{changemargin}{1cm}{1cm}
    Is geography destiny? What is the role of first-nature geography in determining prosperity? This paper estimates the effect of randomly removing and introducing favorable first-nature geography to a specific region using a difference in difference design. In 1825 a storm created a new natural navigable waterway, bringing trade and prosperity to the otherwise relatively isolated northwestern Denmark. 700 years prior, the same event happened in reverse, when a previous channel closed up between 1086 and 1208. The elasticity of geography-induced market access is estimated to be 1.6, corresponding to 26.7 percent population growth within a generation of the event. Demonstrated mechanisms include trade, fertility, fishing, and the rise of manufacturing. The central finding is replicated in reverse in a register of dated archaeological sites. The 1086-1208 closing caused fewer buildings and sites containing coins. The general insight is the same: First-nature geography determines the levels and location of prosperity.

@@

This paper provides evidence of the effects of these shocks in a difference-in-difference framework. This is done using historical census data, a large database of Northern European historical trade (the Sound Toll Register), and the densities of archaeological findings. It is shown that the 1825 breach led to an increase in population of 26.7 percent in 1901. To document the mechanisms, I have collected new data on all Danish occupations (13.5 million observations) in the period 1787-1901. The Danish censuses contain a description of each person's occupation. A new machine learning procedure, \textit{OccCANINE}, is applied to these descriptions to convert them into standardized HISCO codes with high accuracy \citep{leeuwen2002hisco, dahl2024breaking}. The HISCO-coded census data is then used to estimate the effect on occupational structure. The key finding is that the channel caused both fishing and manufacturing to grow. An important question is whether the improved geography caused growth or simply reallocation of prosperity. By using age, gender, and birthplace information it is shown that the population growth came primarily from fertility - not migration. Denmark in the 19th century was in a post-Malthusian regime, where improved living standards cause fertility \citep{Jensen2022, Galor2011}. Taken together, the evidence of this paper suggests that intrinsic growth is the most likely explanation of the observed effect. That is, improved first-nature geography seemingly causes intrinsic growth - not just a reallocation of growth that would have occurred elsewhere.

@@

The improved first-nature geography also caused the population to increase. Figure \ref{fig:pop1} shows regression results from estimating equation \ref{eq:eq501}. Panel (a) shows the simplest estimate, where being affected by the Agger channel is defined as a dummy for being in the West Limfjord. This allows parameters to be interpreted as the approximate percentage increase in the affected region caused by the channel. By 1901 the parishes of the West Limfjord region had a 23.6 log-points (26.7 percent) larger population than what would have been expected by parishes from elsewhere and the population in 1801. Panel (b) shows the results when 'affected' is defined as estimated improved market access from the channel. This shows that by 1901, affected areas had experienced a population growth equivalent to an elasticity of market access of 1.59.\footnote{This is similar to the magnitude found for the Panama Canal by \cite{rauch2022a}.}

@@

The empirical findings of this study reveal that the sudden emergence of the Agger channel in 1834 had a profound and lasting impact on the economic landscape of Denmark's West Limfjord region. Using a difference-in-differences design, it is shown that the channel's creation led to a 26.7 percent increase in the population of affected parishes by 1901 compared to unaffected areas. This population growth was driven primarily by enhanced market access, which facilitated a significant rise in trade, fishing and manufacturing activities. Moreover, the findings indicate that this growth was intrinsic, with higher fertility rates contributing to the increased population rather than migration. Additionally, a reverse natural experiment from the 12th century, when a similar channel closed, resulted in a substantial decline in economic activity, as evidenced by archaeological data. These results underscore the critical role of first-nature geography in driving economic prosperity.
